Interest Flooding Attack and Countermeasures in Named Data Networking

### 2013-01-15 03-interest-flooding.tex
Final pass
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@941 df1a90cd-247b-4364-bdaf-287a704eb499

Since an NDN network fetches data by its name, an adversary cannot target specific routers or end-hosts.
%% as neither is required to have a globally routable network identity/address.
However, an adversary can target a specific namespace.
For example in Fig.~\ref{fig:flooding example}, if the Data producer is the exclusive owner of \ndnName{/foo/bar} namespace, both router B and the Data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches.%
\footnote {This example assumes that the adversary floods the network with unique data names carrying \ndnName{/foo/bar} prefix to make them effective. It also assumes the producer is single-homed and the data is not replicated elsewhere. With multi-homed producers or replicated data, NDN would likely to cope better with DDoS attacks due its native multipath and adaptive forwarding~\cite{adaptive-forwarding}.}
A large volume of such malicious Interests can disrupt service quality in NDN network in two ways: \emph{create network congestion} and \emph{exhaust resources on routers}.

@@

An effective Interest flooding attack is non-trivial to create.
To target a specific namespace (e.g., \ndnName{/newyorktimes/}), an adversary needs to make sure that (1) the expressed Interests are routed as close to the Data producer as possible, and (2) the corresponding PIT entries are stored at intermediate NDN routers for as long as possible.
%In order to achieve both requirements effectively, malicious Interests should avoid Interest collapsing and 
To avoid an Interest being served from caches of intermediate routers, every single malicious Interest has to request an unpopular or non-existing content, e.g., content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damage from each malicious Interest in the network. However, other less effective strains of Interest flooding attacks can also be mitigated by applying the same or similar countermeasures described in the next section.

>>>>

Since an NDN network fetches data by its name, an adversary cannot target specific routers or end-hosts.
%% as neither is required to have a globally routable network identity/address.
However, an adversary can target a specific namespace.
For example in Fig.~\ref{fig:flooding example}, if the data producer is the exclusive owner of \ndnName{/foo/bar} namespace, both router B and the data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches.%
\footnote {This example assumes that the adversary floods the network with unique data names carrying \ndnName{\small /foo/bar} prefix to make them effective. It also assumes the producer is single-homed and the data is not replicated elsewhere. With multi-homed producers or replicated data, NDN would likely to cope better with DDoS attacks due its native multipath and adaptive forwarding~\cite{adaptive-forwarding}.}
A large volume of such malicious Interests can disrupt service quality in NDN network in two ways: \emph{create network congestion} and \emph{exhaust resources on routers}.

@@

An effective Interest flooding attack is non-trivial to create.
To target a specific namespace (e.g., \ndnName{/newyorktimes/}), an adversary needs to make sure that (1) the expressed Interests are routed as close to the data producer as possible, and (2) the corresponding PIT entries are stored at intermediate NDN routers for as long as possible.
%In order to achieve both requirements effectively, malicious Interests should avoid Interest collapsing and 
To avoid an Interest being served from caches of intermediate routers, every single malicious Interest has to request an unpopular or non-existing content, e.g., content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damage from each malicious Interest in the network. However, other less effective strains of Interest flooding attacks can also be mitigated by applying the same or similar countermeasures described in the next section.

### 2013-01-15 04-design.tex
Final pass
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@941 df1a90cd-247b-4364-bdaf-287a704eb499

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers to control the inbound data traffic by controlling the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet-switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of data requested that can fully utilize the downstream link (estimated from the number of forwarded Interests) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens (the pending \emph{Interests Limit}) for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

@@

Fig.~\ref{fig:ratio example} illustrates the resulting dynamics of the statistic during and after an Interest flooding attack. The attack duration is from 10 to 70~seconds. Prior to start of the attack, the percentage of unsatisfied Interests is zero.  
The statistics build up rapidly as soon as Interests start to time out, which happens approximately one second after the start of the attack.%\footnote{Again, we are assuming that Interests are admitted for a maximum period one second.}
For the duration of the attack (10--70~seconds), the percentage of unsatisfied Interests is close to 100\%: 
when the ratio is close to 100\%, routers drop all incoming Interests, resulting in decaying of the statistics until a new Interest is admitted, which eventually brings statistics back near 100\% point.
Finally, the ratio exponentially decays after the attack ceases.

@@

Having successfully implemented a technique to gather statistics on Interest satisfaction ratios, our next challenge is in using this ratio to penalize malicious Interests. A straightforward method to achieve this enforcement is to use the Interest satisfaction ratio as the direct probability for accepting (forwarding) or rejecting an incoming Interest (see Pseudocode~\ref{alg:probabilistic model}).
%Apart from the Interest satisfaction statistics generation, there is a question how this statistics can be used to actually enforce prioritization and penalizing of Interests.

>>>>

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers to control the inbound data traffic by controlling the number of outstanding Interests in the network. 
One simple implementation technique  is for an NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet-switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of data requested that can fully utilize the downstream link (estimated from the number of forwarded Interests) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens (the pending \emph{Interests Limit}) for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:
%
\begin{equation}
\label{eq:interest limit}
\small \mathrm{Interest\ Limit} = Delay\ [s] \cdot \frac{\mathrm{Bandwidth\ [Bytes/s]}}{\mathrm{Data\ packet\ size\ [Bytes]}}
\end{equation}

@@

Fig.~\ref{fig:ratio example} illustrates the resulting dynamics of the statistic during and after an Interest flooding attack. The attack duration is from 10 to 70~seconds. Prior to start of the attack, the percentage of unsatisfied Interests is zero.  
The statistics build up rapidly as soon as Interests start to time out, which happens approximately one second after the start of the attack.
%\footnote{Again, we are assuming that Interests are admitted for a maximum period one second.}
For the duration of the attack (10--70~seconds), the percentage of unsatisfied Interests is close to 100\%: 
when the ratio is close to 100\%, routers drop all incoming Interests, resulting in decaying of the statistics until a new Interest is admitted, which eventually brings statistics back near 100\% point.
Finally, the ratio exponentially decays after the attack ceases.

@@

Having successfully implemented a technique to gather statistics on Interest satisfaction ratios, our next challenge is in using these ratios to penalize malicious Interests. A straightforward method to achieve this enforcement is to use the Interest satisfaction ratio as the direct probability for accepting (forwarding) or rejecting an incoming Interest (see Pseudocode~\ref{alg:probabilistic model}).
%Apart from the Interest satisfaction statistics generation, there is a question how this statistics can be used to actually enforce prioritization and penalizing of Interests.

### 2013-01-15 05-evaluation.tex
Final pass
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@941 df1a90cd-247b-4364-bdaf-287a704eb499

% Traffic pattern
In our experiments, we assumed that legitimate users express Interests at constant average rates with randomized time gap between two consecutive Interests, where the random number for the gap follows a uniform distribution. We believe that this traffic pattern provides a reasonable approximation of traffic mix from all network users without excessive buffering. To quantify the behavior of our mitigation strategies under a worse case attack scenario, we assumed that all the attackers send junk Interests as fast as they can, no Interest---including legitimate ones---can be satisfied from caches.  Also, we configured routers for single-path Interest forwarding and there is only one single-homed producer for the prefix under attack.
%(remember, that Interest limits will not allow real flooding of Intersests in all of the designed mitigation algorithms).

@@

% Fixed parameters
We set the \emph{delay} and \emph{data size} parameters for the Interest limit calculation to an initial %fixed 
value for every node in the simulated topology. In particular, for the small-scale binary tree topology, we set delay to 80~ms, while for the large-scale ISP topology we set it to 330~ms (the order of the largest RTT). The data size is 1100~bytes for all simulation runs and topologies.
% THIS IS BULLSHIT! There is absolutely no logic to set the delay to the same value on all nodes in the network. We discussed multiple times that you should fix it so that the delays are realistic and gets smaller as the nodes gets closer to the producer...

@@

% Topology description
%To assess baseline quality of the designed Interest flooding attack mitigation methods, we evaluated them first using a %simplistic small-scale binary tree topology 
In Fig.~\ref{fig:small-scale}, we depict the binary-tree topology that we used for our initial experiments.
Legitimate users as well as attackers were placed on leaf nodes (red nodes) as show in the figure. There are 16 end users (both legitimate and attackers) in this topology, each expressing Interests that are routed towards a single data producer placed at the root of the tree.  Each link in this topology is assigned a bandwidth of 10~Mbps and a randomized propagation delay ranging from 1 to 10 ms.

@@

The effectiveness of this algorithm stems from the fact that routers can differentiate and limit malicious Interests into the network%, thereby ensuring availability of network resources for serving legitimate users
. The observed periodic dips in the Interest satisfaction ratios of legitimate users in Fig.~\ref{fig:small-scale attack progress} is a direct result of Interest satisfaction rate statistics decaying with time. The 50-second period approximately corresponds to the selected exponential decaying parameter $\alpha=e^{−1.0/30.0}$, which decays statistics to $1/e$ of the initial value that ranges from 30~seconds to 50~seconds. 
%The primary reason that such minimum peaks exist is the fact that 
When Interests from attackers start to get readmitted, they cause degradation of statistics on routers close to the producer (i.e., routers that observe mixed traffic from legitimate and malicious users). Consequently, this degradation reduces the probability of legitimate Interests getting through (see Section~\ref{sec:probabilistic}) until malicious Interests are ``pushed back'' to the edge.

@@

Our next goal is to study the effectiveness of our mitigation algorithms as a function of increasing adversaries in the network.
To this end, we vary the percentage of attackers in the topology from 6\% to over 50\%. Since the total number of end users in the topology is constant, as the number of attackers increases, the number of legitimate users decreases. All other parameters and experimental setup is the same with the previous experiment. As before, for each mitigation algorithm, we perform 10 independent simulation runs.
%The second set of conducted experiments was aimed to answer the question of the effect and quality of the Interest flooding attack mitigation algorithm under different attack volume.
%To do this, for each algorithm we varied the number of adversary nodes in the topology, keeping the total number of client nodes constant: 1 attacker and 15 legitimate users, 3 attackers and 13 legitimate users, etc.

@@

In Fig.\ref{fig:small-scale-topo boxplot} we present  the Interest satisfaction ratio for legitimate users aggregated over the 10 simulation runs for the first minute of the attack. The results are as expected---for all three mitigation algorithms, as the percentage of attackers in the network increases, the lower is the Interest satisfaction ratio for legitimate users.
In the case of token bucket with per-interface fairness algorithm, just 3 attackers can halve the quality of service for the remaining 13 legitimate users. While the two intelligent attack mitigation algorithms also show a decline in service quality as the percentage of attackers increases, this decline is much more gradual and marginal. In the case of satisfaction-based pushback algorithm, over 90\% of Interests from legitimate users are satisfied even if the 50\% of end nodes are malicious.

@@

In Fig.~\ref{fig:large-scale}, we summarize our results aggregated over 10 simulations runs for each mitigation algorithm for the scenario, where the data producer is placed at a gateway node. We observe similar results for the data producer placed at the backbone node as well. As in the case of the binary-tree topology experiments, we observe that the token bucket with per interface fairness is the most ineffective algorithm and satisfaction-based pushback is the most effective one.  Interest satisfaction percentage for legitimate users are approximately 25\% and close to 100\% respectively for these two mitigation methodologies.

@@

Satisfaction-based Interest acceptance algorithm, which was very effective for binary-tree topology, is completely ineffective when deployed in a larger and more realistic topology. For the duration of the attack, legitimate users experience poor quality of service with only 25\% of their Interests being satisfied and continue to experience degraded service long after the attack has stopped. This poor performance, as detailed in Section~\ref{sec:probabilistic}, is due to the fact that each router in the path makes an independent, uncoordinated decision on whether to forward or drop an Interest. In the case of the larger-topology, with much higher average hop count, Interest packets from legitimate users have a very low probability of reaching the data producer, resulting in poor Interest satisfaction statistics, which in turn leads to further penalization of new Interests from them.

All our evaluations leaves us with the conclusion that, among the three techniques we tested under various topology and attacker concentrations, the satisfaction-based pushback is the most effective one in mitigating Interest flooding attacks.% and it restricts malicious Interests from even entering the network. 
%The only short periods of time when malicious Interests are getting admitted to the network is when routers have either no prior knowledge about per-interface satisfaction ratios (the initial period of the attack) or such knowledge becomes stale (statistics decaying during the attack). As soon as the knowledge is obtained or refreshed, the service for legitimate users returns to norm.

>>>>

% Traffic pattern
In our experiments, we assumed that legitimate users express Interests at constant average rates with randomized time gap between two consecutive Interests, where the random number for the gap follows a uniform distribution. We believe that this traffic pattern provides a reasonable approximation of traffic mix from all network users without excessive buffering. To quantify the behavior of our mitigation strategies under a worst-case attack scenario, we assumed that all the attackers send junk Interests as fast as they can, no Interest---including legitimate ones---can be satisfied from caches.  Also, we configured routers for the single-path Interest forwarding and there is only one single-homed producer for the prefix under attack.
%(remember, that Interest limits will not allow real flooding of Intersests in all of the designed mitigation algorithms).

@@

% Fixed parameters
We set the \emph{delay} and \emph{data size} parameters for the Interest limit calculation (\ref{eq:interest limit}) to a fixed 
value for every node in the simulated topology. In particular, for the small-scale binary tree topology, we set delay to 80~ms, while for the large-scale ISP topology we set it to 330~ms (the order of the largest RTT). The data size is 1100~bytes for all simulation runs and topologies.
% THIS IS BULLSHIT! There is absolutely no logic to set the delay to the same value on all nodes in the network. We discussed multiple times that you should fix it so that the delays are realistic and gets smaller as the nodes gets closer to the producer...

@@

% Topology description
%To assess baseline quality of the designed Interest flooding attack mitigation methods, we evaluated them first using a %simplistic small-scale binary tree topology 
In Fig.~\ref{fig:small-scale}, we depict the binary-tree topology that we used for our initial experiments.
Legitimate users as well as attackers were placed on leaf nodes (top row of red nodes) as shown in the figure. There are 16 end users (both legitimate and attackers) in this topology, each expressing Interests that are routed towards a single data producer, placed at the root of the tree.  Each link in this topology is assigned a bandwidth of 10~Mbps and a randomized propagation delay ranging from 1 to 10 ms.

@@

The effectiveness of this algorithm stems from the fact that routers can differentiate and limit malicious Interests into the network%, thereby ensuring availability of network resources for serving legitimate users
. The observed periodic dips in the Interest satisfaction ratios of legitimate users in Fig.~\ref{fig:small-scale attack progress} is a direct result of Interest satisfaction rate statistics decaying with time. The 50-second period approximately corresponds to the selected exponential decaying parameter $\alpha=e^{−1.0/30.0}$, which decays statistics to $1/e$ of the initial value within 30~seconds and to $\approx$20\% within 50~seconds.
 % that ranges from 30~seconds to 50~seconds. 
%The primary reason that such minimum peaks exist is the fact that 
When Interests from attackers start to get readmitted, they cause degradation of statistics on routers close to the producer (i.e., routers that observe mixed traffic from legitimate and malicious users). Consequently, this degradation reduces the probability of legitimate Interests getting through (see Section~\ref{sec:probabilistic}) until malicious Interests are ``pushed back'' again to the edge.

@@

Our next goal is to study the effectiveness of our mitigation algorithms as a function of increasing adversaries in the network.
To this end, we vary the percentage of attackers in the topology from 6\% to over 50\%. Since the total number of end users in the topology is constant, as the number of attackers increases, the number of legitimate users decreases. All other parameters and experimental setup are consistent with the previous experiment. As before, for each mitigation algorithm, we perform 10 independent simulation runs.
%The second set of conducted experiments was aimed to answer the question of the effect and quality of the Interest flooding attack mitigation algorithm under different attack volume.
%To do this, for each algorithm we varied the number of adversary nodes in the topology, keeping the total number of client nodes constant: 1 attacker and 15 legitimate users, 3 attackers and 13 legitimate users, etc.

@@

In Fig.\ref{fig:small-scale-topo boxplot} we present  the Interest satisfaction ratio for legitimate users aggregated over the 10 simulation runs for the first minute of the attack. The results are as expected---for all three mitigation algorithms, as the percentage of attackers in the network increases, the lower is the Interest satisfaction ratio for legitimate users.
In the case of token bucket with per-interface fairness algorithm, just 3 attackers can halve the quality of service for the remaining 13 legitimate users. While the two intelligent attack mitigation algorithms also show a decline in service quality as the percentage of attackers increases, this decline is much more gradual and marginal. In the case of satisfaction-based pushback algorithm, during the first minute of the attack over 90\% of Interests from legitimate users are satisfied even if the 50\% of end nodes are malicious.

@@

In Fig.~\ref{fig:large-scale}, we summarize our results aggregated over all simulations runs for each mitigation algorithm for the scenario, where the data producer is placed at a gateway node. We observe similar results for the data producer placed at the backbone node as well. Unlike the binary-tree topology experiments, we observe that the satisfaction-based pushback is the most effective algorithm, while both the token bucket with per interface fairness and satisfaction-based Interest accept have poor performance.  Interest satisfaction percentage for legitimate users are close to 100\% and approximately 30\% and 25\% respectively for these mitigation methodologies.

@@

Satisfaction-based Interest acceptance algorithm, which was very effective for binary-tree topology, is completely ineffective when deployed in a larger and more realistic topology.  For the duration of the attack, legitimate users experience poor quality of service with only 25\% of their Interests being satisfied and continue to experience degraded service long after the attack has stopped. This poor performance, as detailed in Section~\ref{sec:probabilistic}, is due to the fact that each router on the path makes an independent, uncoordinated decision on whether to forward or drop an Interest. In the case of the larger-topology, with much higher average hop count, Interest packets from legitimate users have a very low probability of reaching the data producer, resulting in poor Interest satisfaction statistics, which in turn leads to further penalization of new Interests from them.

All our evaluations leaves us with the conclusion that, among the three techniques we tested under various topology and attacker concentrations, the satisfaction-based pushback is the most promising one in mitigating Interest flooding attacks.% and it restricts malicious Interests from even entering the network. 
%The only short periods of time when malicious Interests are getting admitted to the network is when routers have either no prior knowledge about per-interface satisfaction ratios (the initial period of the attack) or such knowledge becomes stale (statistics decaying during the attack). As soon as the knowledge is obtained or refreshed, the service for legitimate users returns to norm.

### 2013-01-15 08-conclusion.tex
Small typo
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@940 df1a90cd-247b-4364-bdaf-287a704eb499
Final pass
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@941 df1a90cd-247b-4364-bdaf-287a704eb499

NDN is a newly proposed future Internet architecture, thus it is important to address its resilience in face of DDoS attacks. As a first step in understanding the DDoS threats in NDN, we first examined a specific instance of DDoS attacks, Interest flooding, and the severe service degradation such an attack may cause to legitimate users. We then leverage the key features of the NDN architecture to design, develop and evaluate three mitigation strategies.
We performed detailed simulations on a range of topologies to quantify the effectiveness of our algorithms.   
The most effective algorithm---satisfaction-based pushback---was successful in almost completely shutting down the attackers so that they cause little or no service impact on legitimate users.

>>>>

NDN is a newly proposed future Internet architecture, thus it is important to address its resilience in face of DDoS attacks. As an initial step in understanding the DDoS threats in NDN, we first examined a specific instance of DDoS attacks, Interest flooding, and the severe service degradation such an attack may cause to legitimate users. We then leveraged the key features of the NDN architecture to design, develop, and evaluate three mitigation strategies.
We performed detailed simulations on a range of topologies to quantify the effectiveness of our algorithms.   
The most effective algorithm---satisfaction-based pushback---was successful in almost completely shutting down the attackers, so that they cause little or no service impact on legitimate users.
